---
title: "Create ImmuneSpace Word Frequency List"
output: rmarkdown::html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Create ISFreqsList}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How the list is made
In the `makeContextVector()` function, the utility function `vec2words()` is used.  This function
takes in a character vector and then performs a series of operations:
1. De-duplicate the elements in the vector
2. Remove NA elements
3. Substitute a single whitespace string for certain punctuation
4. Condense any multiple whitespaces into one
5. Split each element into a sub-list of character strings with split = " ", aka words
6. Collect all words into one list
7. Set all letters to lowercase
8. Remove digit-only words
9. For words with only alpha characters, break apart compound words
10. For words with mixed alphanumeric characters, keep only those with "cd|il"
11. Remove words with less than 2 characters

This order of operations is similar to those implemented by the openRefine project by Google. 
The key decisions are described below:
1. De-duplication prior to splitting - By de-duplicating first, we bias the corpus towards words
that are found across studies instead of within studies. This is because it is common for elements
in the same study to have been copy-pasted and therefore not unique.  If there is a misspelling,
then this error is propagated.  Therefore, we seek to mitigate possible copy-paste errors using
`unique()` at this early stage.
2. Removing punctuation and digits - Although some punctuation is used in different terms, such
as "anti-hemoglobin", it was decided to limit punctuation only to biologically-specific terms
such as "CD4+" and "IL-2"

```{r, create-word-lists}
# Create word frequencies list object
ISqueries <- read.csv("ImmuneSpaceQueries.csv", stringsAsFactors = F)
lsWords <- mapply(makeISContextVector, 
                  tableName = ISqueries$tableName,
                  colName = ISqueries$columnName)
names(lsWords) <- paste0(ISqueries$tableName, "-", ISqueries$columnName)
ISFreqsByNm <- sapply(lsWords, table)
save(ISFreqsByNm, file = "../data/ISFreqsByNm.rda")
ISFreqsAll <- table(unlist(lsWords))
save(ISFreqsAll, file = "../data/ISFreqsAll.rda")
```

